---
title: "Basic_R_Proficiency_HW_Nobel"
author: "Nobel"
date: "2026-09-23"
output:
  pdf_document: default
  html_document: default
---

Homework: Basic R Proficiency

# Load Libraries
```{r}
library(tidytext)
library(tidyverse)
```

# Import Data
```{r}
mob_data <- rio::import("https://raw.githubusercontent.com/wellsdata/CompText_Jour/main/data/mob_analysis_data.csv")
```

My memo: 

# Using code, describe the number of rows and columns in the data set

```{r}
nrow(mob_data)
ncol(mob_data)
```
# There are 3836 rows and 7 columns in the data set


# Create a table that displays a count of the top 15 newspaper entries
```{r}
Top_15_Newspaper_Entries <- mob_data |>
  group_by(Newspaper) |> 
  count(Newspaper) |>
  arrange(desc(n)) |>
  head(15)
Top_15_Newspaper_Entries
```
# A tibble:15 × 2
Groups:Newspaper [15]
Newspaper
<chr>
n
<int>
NA	336			
The Birmingham age-herald (Birmingham, Ala...)	118			
The daily worker (Chicago, Ill.)	64			
The Topeka state journal (Topeka, Kansas)	63			
The sun (New York [N.Y.]) 1833-1916	50			
The Seattle post-intelligencer (Seattle, Wash. ...)	49			
The San Francisco call (San Francisco [...)	46			
The morning news (Savannah, Ga.)	40			
Los Angeles herald (Los Angeles [Calif.]) 1900-1911	37			
Watertown republican (Watertown, Wis.)	36			
River Falls journal (River Falls, Pi...)	34			
Vernon County censor (Viroqua, Wis.)	30			
The Iola register (Iola, Allen Cou...)	29			
The Daily Ardmoreite (Ardmore, Okla.)	28			
The Salt Lake herald (Salt Lake City ...)	28			
#15 rows		

# Create a table that counts the different entries in the "keyword" column

```{r}
Keywords <- mob_data |> 
  group_by(keyword) |>
  count(keyword) |> 
  arrange(desc(n))
Keywords
```
# There are 37 different entries in the keyword column; mob, masked and mobs are the top entries.


# Create a table that counts all specific articles by year

```{r}
Years <- mob_data %>% 
  group_by(Year) %>% 
  count(Year) %>% 
  arrange(desc(n))
Years
```
# There are 110 different years in the data set. The top years by article count are 1903 with 246 articles, 1906 with 207 articles, and 1901 with 197 articles. 	

# Create a simple column ggplot chart that shows the total entries by year

```{r}
ggplot(Years,aes(x = Year, y = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  labs(title = "Total Entries of Articles by Year",
       subtitle = "Mob Lynching Article Data, 1805-1955",
       caption = "Graphic by Shahariar Nobel, 9-26-2026",
       y="Number of Articles",
       x="Year")
```
# The memo
Based on the analysis of the mob lynching data set, it contains 3,836 rows and 7 columns. The newspaper columns show a collection of articles coming from many different publications, although the number of articles by newspapers is uneven. An inquiry of the top 15 newspapers by entries showed that The Birmingham Age-Herald has the largest number of identified entries with 118, followed by The Daily Worker with 64 and The Topeka State Journal with 63 in the data set. However, the highest 336 entries were under an unidentified newspaper name (NA), which suggests that a notable portion of the articles are missing their publication information. The analysis of the keyword column returned 37 different keywords. Among them, “mob” appears 3,273 times. The next most common keyword is “masked,” which appears 158 times, and “mobs,” which appears 96 times. This suggests that the keywords are heavily concentrated around the term “mob”. The count by year shows 110 different years, with the largest count found in 1903 with 246 articles, followed by 1906 with 207 and 1901 with 197 articles. A graph of the years shows the article count rising sharply during the decade from 1890 to the 1990s. From the analysis, some questions remain about the data set. Why are the newspaper names missing for those 336 articles?  Also, why is the dataset missing articles from some years that show no values in the graph?




