---
title: "Basic_R_Proficiency_HW_Nobel"
author: "Nobel"
date: "2026-09-23"
output: html_document
---

Homework: Basic R Proficiency

# Load Libraries
```{r}
library(tidytext)
library(tidyverse)
```

# Import Data
```{r}
mob_data <- rio::import("https://raw.githubusercontent.com/wellsdata/CompText_Jour/main/data/mob_analysis_data.csv")
```

My memo: 

# Using code, describe the number of rows and columns in the data set

```{r}
nrow(mob_data)
ncol(mob_data)
```
# There are 3836 rows and 7 columns in the data set


# Create a table that displays a count of the top 15 newspaper entries
```{r}
Top_15_Newspaper_Entries <- mob_data |>
  group_by(Newspaper) |> 
  count(Newspaper) |>
  arrange(desc(n)) |>
  head(15)
Top_15_Newspaper_Entries
```
# A tibble:15 × 2
Groups:Newspaper [15]
Newspaper
<chr>
n
<int>
NA	336			
The Birmingham age-herald (Birmingham, Ala...)	118			
The daily worker (Chicago, Ill.)	64			
The Topeka state journal (Topeka, Kansas)	63			
The sun (New York [N.Y.]) 1833-1916	50			
The Seattle post-intelligencer (Seattle, Wash. ...)	49			
The San Francisco call (San Francisco [...)	46			
The morning news (Savannah, Ga.)	40			
Los Angeles herald (Los Angeles [Calif.]) 1900-1911	37			
Watertown republican (Watertown, Wis.)	36			
River Falls journal (River Falls, Pi...)	34			
Vernon County censor (Viroqua, Wis.)	30			
The Iola register (Iola, Allen Cou...)	29			
The Daily Ardmoreite (Ardmore, Okla.)	28			
The Salt Lake herald (Salt Lake City ...)	28			
#15 rows		

# Create a table that counts the different entries in the "keyword" column

```{r}
Keywords <- mob_data |> 
  group_by(keyword) |>
  count(keyword) |> 
  arrange(desc(n))
Keywords
```
# There are 37 different entries in the keyword column; mob, masked and mobs are the top entries.


# Create a table that counts all specific articles by year

```{r}
Years <- mob_data %>% 
  group_by(Year) %>% 
  count(Year) %>% 
  arrange(desc(n))
Years
```
# There are 110 different years in the data set. The top years by article count are 1903 with 246 articles, 1906 with 207 articles, and 1901 with 197 articles. 	

# Create a simple column ggplot chart that shows the total entries by year

```{r}
ggplot(Years,aes(x = Year, y = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  labs(title = "Total Entries of Articles by Year",
       subtitle = "Mob Lynching Article Data, 1805-1955",
       caption = "Graphic by Shahariar Nobel, 9-26-2026",
       y="Number of Articles",
       x="Year")
```


