---
title: "Metadata"
author: "Wells"
date: "2026-09-15"
output: html_document
---

Class exercise to import your project metadata from the spreadsheet into R 

--Create a folder called "class_project" in your code folder 

--Save this file to the "class_project" folder 

--Move your Nexis spreadsheet to the folder 

--Create a project file on R Studio: File \| New Project \| \*\*\* select Existing folder \| "class_project" 

--After project.Rproj launches, then launch this file, metadata.Rmd, and complete the tutorial

**By doing this, you avoid the ring of hell where R can't find your data**

#load tidyverse, janitor, rio

```{r}

library(tidyverse)
library(janitor)
library(rio)
install.packages("readxl")
library(readxl)

```

#Import your data

```{r}

data <- read_excel("SKN_Trust_Data.XLSX") |> 
  #clean_names()
View(data)

```

# Clean names

```{r}
data <- data |> 
  clean_names()

```

# check data types
```{r}
glimpse(data)
```


Next Steps: Count years, and create a ggplot

# Transform date into years. See tutorial
```{r}

data2 <- data |> 
  mutate(year = year(published_date))

```

#year
```{r}
data2 |> 
  count(year)
```

#Pulication
```{r}
data2 |> 
  count(publication_4) |> 
  arrange(desc(n))
```


#publication_location
```{r}
data2 |> 
  count(publication_location) |> 
  arrange(desc(n))
```

#length check
```{r}
#crate a numeric ength column

data_lenghth <- data2 |> 
  mutate(length_num = as.numeric(length)) |>
  filter()
```



# Create a ggplot
```{r}
data2 |> 
  count(year) |>
   ggplot(aes(x = year, y = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  scale_x_continuous(breaks=c(2021:2026)) + #fixes wonky x axis
  labs(title = "Years for Coverage of Trust in Media",
       subtitle = "Exercise NexisUni Run",
       caption = "Graphic by Shahariar Nobel, 9-16-2026",
       y="Number of Calls",
       x="Year")



```

#Count publications

```{r}
data2 |> 
  count(publication_4)
```


#Visualize publications

```{r}
data2 |> 
  count(publication_4) |> 
#ggplot(aes(x = publication_4, y = n,
             #fill = n)) +
  ggplot(aes(x = reorder(publication_4, n), y = n, fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  #scale_x_continuous(breaks=c(2021:2026)) + #fixes wonky x axis
  coord_flip() +
  labs(title = "Publications Covered Trust in Media",
       subtitle = "Exercise NexisUni Run",
       caption = "Graphic by Shahariar Nobel, 9-16-2026",
       y="Number of Articles",
       x="Year")
```




# NOTES ON FIXING COMPLEX Date COLUMN WITH HOURS AND MINUTES

```{r}
women <- women |> 
  mutate(date = as.Date(strptime(published_date, format = "%B %d, %Y %A")))
```